---
title: "Conditional Loops Activity"
author: "Lewis White"
date: "2022-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

TASK 1

Create an object called pm2_5 with a value of 48 (representing Particulate Matter 2.5, an indicator for air quality, in μgm3 (see more about PM2.5 here).

Write an if - else if - else statement that returns “Low to moderate risk” if pm2_5 (for Particulate Matter 2.5) is less than 100, “Unhealthy for sensitive groups” if PM 2.5 is 100 <= pm2_5 < 150, and “Health risk present” if PM 2.5 is >= 150.

Test by changing the value of your pm2_5 object and re-running your statement to check.

```{r}
pm2_5 <- 48

if (pm2_5 < 100) {
  print("Low to moderate risk")
} else if (pm2_5 >= 100 & pm2_5 < 150) {
  print("Unhealthy for sensitive groups")
} else {
  print("Health risk present")
}
```


TASK 2

Store the string “blue whale” as an object called species. Write an if statement that returns “You found a whale!” if the string “whale” is detected in species, otherwise return nothing. Test by changing the species string & re-running to see output.


```{r}
species <- c("blue whale")

if (str_detect(species, "whale")) {
  print ("You found a whale")
}
```

TASK 3 

Create a vector stored as max_airtemp_c with a value of 24.1. Write an if else statement that will print “Temperature too high” if max_airtemp_c is greater than 27, or “Temperature OK” if temperature is less than or equal to 27.

```{r}
max_airtemp_c <- 24.1

if (max_airtemp_c > 27) {
  print("Temperature too high")
} else {
  print("Temperature OK")
}
```

TASK 4 

Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredent with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:

A veggie burrito is the cost of a base burrito
A chicken burrito costs 3.00 more than a base burrito
A steak burrito costs 3.25 more than a base burrito

```{r}
base_burrito <- 6.50
main_ingredient <- "steak"

if (main_ingredient == "veggie") {
  print(base_burrito)
} else if (main_ingredient == "chicken") {
  print(base_burrito + 3)
} else if (main_ingredient == "steak") {
  print(base_burrito + 3.25)
}
```



TASK 5

Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}
fish <- c(8, 10, 12, 23)
fish_type <- c("goldfish", "tetras", "guppies", "mollies")

fish_prop <- vector(mode = "numeric", length = length(fish))

for (i in seq_along(fish)) {
  prop = fish[i]/sum(fish)
  fish_prop[i] = prop
}
```

TASK 6

There is an existing vector in R called month.name that contains all month names (just try running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).


```{r}
for (i in seq_along(month.name)) {
  print(paste(month.name[i], "is month", i))
}
```

